---
title: "Data Import"
output: github_document
---

```{r setup, echo = FALSE, messsage = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

This document will show how to import data by relative path.Tabbed autocompletion can be used to find the data file. The only function I will use in `janitor` package is `clean names`, so i do not want to use library to load janitor package.

## Import the FAS Litters CSV
```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```


 
### Look at the dataset

Printing in the console.

```{r}
litters_df
```

Look at the first 6 rows and the tail of the dataset.

```{r}
head(litters_df)
tail(litters_df, 10)
```

Skim is the only function i will use in `skimr` package.

```{r}
skimr::skim(litters_df)
```


```{r eval = FALSE}
view(litters_df)
```


## Import FAS Pups

Use relative paths:

```{r}
pups_df = read.csv("data/FAS_pups.csv")

pups_df = janitor::clean_names(pups_df)

pups_df
```



### Look at read_csv options

skipping the first two row (since most data file has data dictionary in the first two rows) and use `col_names = FALSE` make up colum names 

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv"
  )
```

What about missing data.

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", ".")
  )

litters_df = janitor::clean_names(litters_df)

pull(litters_df, gd0_weight)

```

What if we code `group` as a factor variable?

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group = col_factor()
    )
  )

litters_df
```



## Import an excel file

Import MLB 2011 summary data and use `sheet=` to choose sheet if there are many sheets. `Range=` specify the range of columns and rows using Excel syntax (like B6 to G18).

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet = "mlb11")
```


## Import SAS data

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")

pulse_df
```


## Never use read.csv()

Never do this either: dollar sign

```{r}
litters_df$L
```





