---
title: "Data Manipulation"
output: github_document
---

```{r setup, echo = FALSE, messsage = FALSE}
library(tidyverse)
```

This document will show how to *manipulate* data.

Import the two datastes that we're going to manipulate.

```{r}
litters_df = 
  read_csv(file = "data/FAS_litters.csv", na = c("NA", "", "."))
litters_df = janitor::clean_names(litters_df)

pups_df = read.csv(file = "data/FAS_pups.csv", na = c("NA", "", "."))
pups_df = janitor::clean_names(pups_df)
```


## `select`

Use `select()` to select some variables

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

Use : to specify a range of column

```{r}
select(litters_df, group:gd18_weight)
```

```{r}
select(litters_df, -litter_number)
```

Renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Select helpers

Keep everything start with `gd`

```{r}
select(litters_df, starts_with("gd"))
```

Remove `litter_number`, but keep everthing else

```{r}
select(litters_df, litter_number, everything())
```

Relocate this column to the beginning

```{r}
relocate(litters_df, litter_number)
```



## `filter`

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

`==` sign is a test for equality

```{r}
filter(litters_df, gd_of_birth == 20)
```


```{r}
filter(litters_df, !(gd_of_birth = 20))
filter(litters_df, gd_of_birth != 20)
```

```{r}
filter(litters_df, gd0_weight >= 22,gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Mod8")
```

Where group is either `of these option`Mod8` or `Con7`

```{r}
filter(litters_df, group %in% c("Mod8", "Con7"))
```

drop_na(litters_df) will remove any row with a missing value
drop_na(litters_df, wt_increase) will remove rows for which wt_increase is missing.


## `mutate`

create new variables and convert group to lower case

```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```


## `arrange`

first in order of `pups_born_alive`, then within each level, in order of gd0_weight

```{r}
arrange(litters_df, pups_born_alive)
```


## `%>%`

```{r}
litters_data_raw = read.csv("data/FAS_litters.csv", na = c("NA", "", "."))
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

Use Pipe operator instead!

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", "", ".")) |> 
  janitor::clean_names() |> 
  select(-pups_survive) |> 
  mutate(wt_gain = gd18_weight - gd0_weight) |> 
  drop_na(gd0_weight)
```











